\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
big push

\abstract{

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing color-dual
  numerators at one-loop directly from Feynman rules. Consideration of
  this theory highlights that the lifting from one-loop
  color-kinematics dual numerators to two loops is just as challenging
  as the progression from tree-level to one-loop.  As a first step in
  understanding the two-loop challenge, we construct color-dual
  two-loop four-point nonlinear sigma model (NLSM) numerators using a
  $D$-dimensional color-dual bootstrap.  The double-copy of the
  constructed NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known
  physical unitarity cuts of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These
  NLSM numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. Using our constructive bootstrap \ace{I am
    really hesitant to call the bootstrap constructive.  While it
    technically constructs, ``constructive'' has some additional
    connotations}, we also find that the most general \emph{local}
  two-loop four-point numerator for pure Yang-Mills fails to satisfy
  global color-kinematics duality. We pinpoint the failure to the
  bow-tie unitarity cut, and discuss a path towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

@@

However, despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our
understanding of the kinematic algebra to the multi-loop level.

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM,
  $\phi^3$' which was redundant.  I don't mind if we say it here
  because I want people to tell people that these are the kinds of
  theories we're thinking about.  We could omit it but then we'd have
  to say that the theory has at least one color factor and is probably
  an adjoint type theory and at that point I'd rather just give an
  example.} -- can be written as
\begin{equation}
  \mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
  \label{eq:gen-amp}
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over only purely cubic graphs $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu,
    Carrasco:2023vjg, Carrasco:2020ywq}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

@@

Many of the 1-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$
terms, indicating that there are additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental BCJ identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half ladder because any
diagram can be reduced to a half ladder through successive
applications of the Jacobi identity.  At 1-loop, every diagram can be
related to one topology as well, in this case the $n$-gon.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction.  The theory will be color dual if the
cubic vertex is antisymmetric and satisfies the Jacobi identity,
specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}P_{b_1 c_1}V^{c_1 a_3b_2}\cdots P_{b_{n-3}c_{n-3}}V^{c_{n-3} a_{n-1}a_n}\,.
\end{equation}
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See ref \cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the loop-level
amplitude integrand must be constructed from an ansatz.  The four
conditions we impose on an integrand ansatz are summarized here and
elaborated on below.\footnote{The four ansatz conditions are similar
  to those of Ref.~\cite{Bern:2015ooa} except that there the authors
  demand that the numerators have the same loop power counting as
  Feynman gauge Feynman rules.  This is an additional assumption that
  is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity }: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections
include examples from a 4-point 2-loop integrand for some generic
color-dual theory like YM.  This is the simplest case that
demonstrates the full machinery of the color-dual bootstrap.  All tree
and 1-loop processes are generated by single basis diagrams (the half
ladder and $n$-gon respectively) so they lack examples of three-term
``boundary'' Jacobi relations to be described shortly.
Choosing the 4-point 2-loop integrand also has the advantage that this
process appears prominently in this paper.

@@

\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH
}
\caption{The 14 non-pathological 4-point 2-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}

Color-kinematics duality is a statement about color factors, kinematic
numerators, and the accompanying Jacobi relations.  Adjoint-type color
factors are associated with cubic graphs, so satisfying
color-kinematics duality implicitly requires expressing the integrand
in terms of such graphs. As mentioned earlier, this is possible to do
even for a theory with quartic (or higher multiplicity) vertices by
multiplying and dividing by propagators.  In the case of the 4-point
2-loop example promised above, there are 14 cubic graph topologies
(shown in \cref{fig:MaxCuts}) ignoring tadpoles and bubble on external
leg (BEL) graphs.

With the integrand expressed in terms of cubic graphs, it is possible
to discuss the main ingredient in color-kinematics duality -- the
Jacobi relations.  The color factors in an integrand obey a set of
Jacobi identities.  Color-kinematics duality states that there
\emph{exists} a way of writing the kinematic numerators of the
integrand so that they obey the same set of Jacobi identities.  The
Jacobi identities can be divided into two categories: \emph{defining}
relations and \emph{boundary} relations.  Defining Jacobi relations
can be used to express any graph in terms of a basis.  For example,
the double box, crossed box, and penta-triangle are related to each
other as follows \be % [[11]]
\label{eq:TwoLoopJacobiOne}
\JacobiOneDoubleBox + \JacobiOneCrossedBox + \JacobiOnePentaTriangle
=0\,,
\end{equation}
where the shared propagator is shown in red.  Using this relation the
crossed box can be written (or \emph{defined}) in terms of the other
two.  In this way, all of the 14 cubic 4-point 2-loop topologies can be
written in terms of a basis of the double box and penta-triangle.
(Any two of the double box, crossed box, and penta-triangle would
suffice for a basis but the planar basis is chosen for simplicity in
this work.)  The numerator of each of the two basis graphs receives
its own ansatz and the numerator of any other graph can be expressed
in terms of these two.

@@

\iffalse
The essence of generalized unitarity is that when internal lines are
placed on-shell, the integrand should factorize into a product of
on-shell tree amplitudes.  The cut to be performed can, itself, be
encoded in a graph.  For example, potential 4-point 2-loop cuts include
\begin{equation}
\label{eq:PotentialCuts}
\text{(a)} \SubtleCut ~~~~ \text{(b)} \PhysicalCutOne{}{}{}{} ~~~~ \text{(c)} \PhysicalCutTwo{}{}{}{} ~~~~ \text{(d)}  \MaxCut
\end{equation}
as well as many more.
In these graphs, all exposed propagators are understood to be cut/on-shell.
As an equation, a unitarity cut looks schematically like
\begin{equation}
\label{eq:SchematicCutEquation}
\sum \limits_\text{topologies} \frac{n \vert_\text{cut}}{\prod \limits_\text{uncut} k^2} = \sum \limits_\text{states} \prod \limits_\text{blobs} A_\text{tree} \, .
\end{equation}
The right hand side of the equation is a product of on-shell tree
amplitudes, represented graphically in \cref{eq:PotentialCuts} by grey
blobs.  States crossing (or propagating across) the cut must be summed
over.  For scalars this sum is trivial but for spinning particles this
is no longer the case and the physical state projector must be
inserted (see Dimitrius paper for simplifying this operation).  The
graph topologies appearing in the left hand side of
\cref{eq:SchematicCutEquation} come from blowing up the grey blobs
into cubic tree graphs in every conceivable way.  The numerator of
each graph is related back to the basis numerators via Jacobi
relations.  For each of these topologies, the numerator of the
associated graph is subjected to the cut constraints and is then
divided by the uncut propagators.  The right hand side is assumed to
be known, which places a constraint on the numerator ansatz in the
left hand side.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop
numerators at any multiplicity for pions.  However, these methods
cannot be reapplied to generate higher-loop numerators, which begs the
question ``is it possible to find globally color-dual numerators
beyond 1-loop?''  We will tackle this question for 4-point 2-loop
scattering in NLSM before moving on to the same process in YM.

>>>>

\abstract{ We make progress in understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators at one-loop directly from Feynman rules. Consideration of
  this theory highlights that the lifting from one-loop
  color-kinematics dual numerators to two loops is just as challenging
  as the progression from tree-level to one-loop.  As a first step in
  understanding the two-loop challenge, we construct color-dual
  two-loop four-point nonlinear sigma model (NLSM) numerators using a
  $D$-dimensional color-dual bootstrap.  The double-copy of the
  constructed NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known
  physical unitarity cuts of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These
  NLSM numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. Using our loop-level bootstrap, we also find that the most general \emph{local} ansatz
  for two-loop four-point pure Yang-Mills fails to satisfy
  global color-kinematics duality. We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

@@

However, despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv,
  Carrasco:2021otn}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies for advancing our
understanding of the kinematic algebra toward the multi-loop level.

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM,
  $\phi^3$' which was redundant.  I don't mind if we say it here
  because I want people to tell people that these are the kinds of
  theories we're thinking about.  We could omit it but then we'd have
  to say that the theory has at least one color factor and is probably
  an adjoint type theory and at that point I'd rather just give an
  example.} -- can be written as
\begin{equation}
  \mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
  \label{eq:gen-amp}
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over the set of purely cubic graphs, $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

@@

Many of the one-loop results presented in \cref{sec:cubic} are informed
by the tree-level double-copy so we will review the tree case here.
The color structure of tree-level amplitudes is particularly simple.
A fully color dressed tree amplitude $\mathcal{A}_n$ can be decomposed
in terms of the trace basis
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-1}} \text{Tr}(T^{a_{\sigma_1}} T^{a_{\sigma_2}}...T^{a_{\sigma_{n-1}}} T^{a_n}) A_n[\sigma_1, \sigma_2,...\sigma_{n-1}, n],
  \label{eq:trace-basis}
\end{equation}
where $T^a$ are generators of the gauge group and, because of the
cyclicity of the trace, one of the legs is held fixed so that the sum
runs over the $(n-1)!$ permutations of the remaining external legs.
The coefficients of the color factors are the color-ordered partial
amplitudes $A_n[...]$.  Any tree-level color factor can be converted
into a linear combination of Del Duca-Dixon-Maltoni (DDM) half-ladder
color factors
\begin{equation}
\text{DDM}[a_1, a_2,... a_n] \equiv f^{a_1 a_2 b_1} f^{b_1 a_3 b_2} f^{b_2 a_4 b_3}...f^{b_{n-3} a_{n-1} a_n}
\end{equation}
associated with the graph
\begin{equation}
\halfLadder
\end{equation}
by repeated application of the Jacobi identity \cite{DixonMaltoni}.
The fully color-dressed tree amplitude can then be re-expressed as
\begin{equation}
  \mathcal{A}_n = \sum \limits_{\sigma\in S_{n-2}} \text{DDM}[a_1, a_{\sigma_2}, a_{\sigma_3}, a_{\sigma_{n-1}}, a_n]  A_n[1, \sigma_2,...\sigma_{n-1}, n]
  \label{eq:ddm-basis}
\end{equation}
with legs $1$ and $n$ fixed so that the sum only contains $(n-2)!$
terms, indicating that there are additional relations -- the
Kleiss-Kuijf (KK) relations \cite{Kleiss:1988ne} -- amongst the
partial amplitudes appearing in the right-hand side of
\cref{eq:trace-basis}.  The KK relations are generic to any theory
with purely adjoint particles.  Color-kinematics duality further
implies that the KK basis is overcomplete and can be reduced to a
basis of $(n-3)!$ amplitudes via the fundamental BCJ identities
\cite{BCJ,Feng:2010my}
\begin{equation}
\sum \limits_{i=2}^{n-1} k_1 \cdot (k_2+...+k_i) A_n[2,...,i,1,i+1,...,n] =0.
\end{equation}
In terms of constructing explicit color-dual solutions at tree level,
it is enough to specify the numerator of the half ladder because any
diagram can be reduced to a half ladder through successive
applications of the Jacobi identity.  At 1-loop, every diagram can be
related to one topology as well, in this case the $n$-gon.  Two-loop
calculations play an important role in our understanding of
color-kinematics duality because this is the first order where
multiple basis graphs are required, at least for a generic theory. We refer the reader to \cite{BCJreview} for more background on the topic.

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}\label{cubicCKLag}
\mathcal{L} = P_V^{ab}P_W^{\mu\nu}\mathcal{O}^a_\mu \Box \mathcal{O}^b_\nu +  V^{abc} W^{\mu \nu \rho}\mathcal{O}^a_\mu \mathcal{O}^b_\nu \mathcal{O}^c_\rho
\end{equation}
where $V^{abc}$ and $W^{\mu \nu \rho}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_\mu$ indexed by quantum numbers $a$ and
$\mu$. The Feynman rules for this theory are simply:
\be\label{CKFrules}
\begin{split}
\propFrule{$a,\mu$}{$b,\nu$}{$k \rightarrow $}{} &= \frac{i}{k^2} P_V^{ab}P_W^{\mu\nu}
\\
\cubicFrule{$a,\mu$}{$b,\nu$}{$c,\rho$}{}{}{} &= i V^{abc}W^{\mu \nu \rho}
\end{split}
\ee
We have introduced $P_V^{ab}$ and $P_W^{\mu\nu}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction, as is the case for $DF^2$ theory \cite{Johansson:2017srf}. Though generally these quadratic dressings are simply delta functions, $P_V^{ab} = \delta^{ab}$, or flat space metrics, $P_W^{\mu\nu}=\eta^{\mu\nu}$. As written, the theory will be color dual if the off-shell cubic vertex is antisymmetric and satisfies the Jacobi identity,
specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where using the Feynman rules of \cref{CKFrules}, the bracketed expression is the half-ladder numerator of the following diagram,
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv N\left[\halfLadder\right]\,,
\end{equation}
and likewise for the $W^{\mu \nu \rho}$ dressing. Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional operators \cite{Carrasco:2022lbm,Carrasco:2022sck}, sometimes including additional propagating fields not originally considered in the spectrum. See ref.~\cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic two-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the loop-level
amplitude integrand must be constructed from an ansatz.  The four
conditions we impose on an integrand ansatz are summarized here and
elaborated on below.\footnote{The four ansatz conditions are similar
  to those of Ref.~\cite{Bern:2015ooa} except that there the authors
  demand that the numerators have the same loop power counting as
  Feynman gauge Feynman rules.  This is an additional assumption that
  is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity}: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

In order to clarify general statements, the following subsections
include examples from a four-point two-loop integrand for some generic
color-dual theory like YM.  This is the simplest case that
demonstrates the full machinery of the color-dual bootstrap.  All tree
and one-loop processes are generated by single basis diagrams (the half
ladder and $n$-gon respectively) so they lack examples of three-term
``boundary'' Jacobi relations to be described shortly.
Choosing the four-point two-loop integrand also has the advantage that this
process appears prominently in this paper.

@@

\paragraph{Color-kinematics duality}
\begin{figure}
\centering
{\MCutC, \MCutD, \sumoCut, \MCutJ,  
 \\
  \doubleBoxCut,  \crossBoxCut,\pentaTriangleCut,  \MCutF, 
\\
\MCutI,  \MCutE, \MCutK, 
\\
 \MCutL, \MCutG,  \MCutH
}
\caption{The 14 non-pathological four-point two-loop cubic diagrams,
  including all dangling tree diagrams.} \label{fig:MaxCuts}
  \end{figure}

Color-kinematics duality is a statement about color factors, kinematic
numerators, and the accompanying Jacobi relations.  Adjoint-type color
factors are associated with cubic graphs, so satisfying
color-kinematics duality implicitly requires expressing the integrand
in terms of such graphs. As mentioned earlier, this is possible to do
even for a theory with quartic (or higher multiplicity) vertices by
multiplying and dividing by propagators.  In the case of the four-point
two-loop example promised above, there are 14 cubic graph topologies
(shown in \cref{fig:MaxCuts}) ignoring tadpoles and bubble on external
leg (BEL) graphs.

With the integrand expressed in terms of cubic graphs, it is possible
to discuss the main ingredient in color-kinematics duality -- the
Jacobi relations.  The color factors in an integrand obey a set of
Jacobi identities.  Color-kinematics duality states that there
\emph{exists} a way of writing the kinematic numerators of the
integrand so that they obey the same set of Jacobi identities.  The
Jacobi identities can be divided into two categories: \emph{defining}
relations and \emph{boundary} relations.  Defining Jacobi relations
can be used to express any graph in terms of a basis.  For example,
the double box, crossed box, and penta-triangle are related to each
other as follows \be % [[11]]
\label{eq:TwoLoopJacobiOne}
\JacobiOneDoubleBox + \JacobiOneCrossedBox + \JacobiOnePentaTriangle
=0\,,
\end{equation}
where the shared propagator is shown in red.  Using this relation the
crossed box can be written (or \emph{defined}) in terms of the other
two.  In this way, all of the 14 cubic four-point two-loop topologies can be
written in terms of a basis of the double box and penta-triangle.
(Any two of the double box, crossed box, and penta-triangle would
suffice for a basis but the planar basis is chosen for simplicity in
this work.)  The numerator of each of the two basis graphs receives
its own ansatz and the numerator of any other graph can be expressed
in terms of these two.

@@

\iffalse
The essence of generalized unitarity is that when internal lines are
placed on-shell, the integrand should factorize into a product of
on-shell tree amplitudes.  The cut to be performed can, itself, be
encoded in a graph.  For example, potential four-point two-loop cuts include
\begin{equation}
\label{eq:PotentialCuts}
\text{(a)} \SubtleCut ~~~~ \text{(b)} \PhysicalCutOne{}{}{}{} ~~~~ \text{(c)} \PhysicalCutTwo{}{}{}{} ~~~~ \text{(d)}  \MaxCut
\end{equation}
as well as many more.
In these graphs, all exposed propagators are understood to be cut/on-shell.
As an equation, a unitarity cut looks schematically like
\begin{equation}
\label{eq:SchematicCutEquation}
\sum \limits_\text{topologies} \frac{n \vert_\text{cut}}{\prod \limits_\text{uncut} k^2} = \sum \limits_\text{states} \prod \limits_\text{blobs} A_\text{tree} \, .
\end{equation}
The right hand side of the equation is a product of on-shell tree
amplitudes, represented graphically in \cref{eq:PotentialCuts} by grey
blobs.  States crossing (or propagating across) the cut must be summed
over.  For scalars this sum is trivial but for spinning particles this
is no longer the case and the physical state projector must be
inserted (see Dimitrius paper for simplifying this operation).  The
graph topologies appearing in the left hand side of
\cref{eq:SchematicCutEquation} come from blowing up the grey blobs
into cubic tree graphs in every conceivable way.  The numerator of
each graph is related back to the basis numerators via Jacobi
relations.  For each of these topologies, the numerator of the
associated graph is subjected to the cut constraints and is then
divided by the uncut propagators.  The right hand side is assumed to
be known, which places a constraint on the numerator ansatz in the
left hand side.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop
numerators at any multiplicity for pions.  However, these methods
cannot be reapplied to generate higher-loop numerators, which begs the
question ``is it possible to find globally color-dual numerators
beyond 1-loop?''  We will tackle this question for four-point two-loop
scattering in NLSM before moving on to the same process in YM.

Ignoring tadpoles and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of
these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to equations with sixteen thousand parameters so a
  different method is desirable.  One possibility is to use a solver
  designed for integral reduction.  However, these solvers typically
  assume that the matrix has a very low density which is not the case
  for equations coming from NLSM which typically have a density of a
  few percent.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such
  separation of sectors in a scalar theory where any product of
  momenta can interfere with any other.  The NLSM constraint equations
  were solved using a custom solver, \texttt{FiniteFieldSolve},
  designed to solve large linear systems of arbitrary density.
  \texttt{FiniteFieldSolve} will be released publicly shortly.\ace{I
    want to rephrase this slightly, but need to think about it a bit
    more}} Imposing the boundary Jacobi relations eliminates all but
4,473 of the original free parameters.  Further imposing graph
symmetries reduces the number of free parameters to 1,243.
